---
title: "Project Title Here"
date: "October 16, 2003"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

### Team Members  
* Simin Liu   
* Iris Chen  
* Latonya Smith  

## Problem Statement

What is the problem you are trying to solve? Why is it important or useful to solve this problem?


## Data Description

Describe the dataset in detail. What is the source? Include URLs. How was it collected? What are the variables? What are the characteristics of the variables?

Run the following code on your data. Be sure to change `eval` to `TRUE` and `mydata` to the name you give your data`.

```{r echo=TRUE, eval=FALSE}
library(readr)
mydata <- read_csv("mydata.csv")


library(qacBase)
df_plot(mydata)
contents(mydata)
histograms(mydata)
barcharts(mydata)
cor_plot(mydata)
```



## Data Preprocessing

Describe any variable transformations, treatment of missing values, recoding and any other data manipulations you believe you will need to do before you can build models.